Ambient: overview of the workflow and design basics

### 2011-08-15 doc/paper_ambient/paper_ambient.tex
some corrections to abstract parallel types

In order to turn Ambient into general purpose framework, we have decoupled the Ambient data representation from the user datatypes. There are two aspects to the parallel datatypes.

First is the object's lifetime - as we employ the technique of accumulated execution - we have to unsure that the objects that were created on the stack are still alive at the time of the actual execution. For this purpose we introduced livelong class that forces derived data types to dplicate themselves in heap memory upon construction if they were allocated on stack. Using this duplicate we can ensure that the object will stay alive until the moment of the actual kernels execution.

Second aspect is the actual connection with the Ambient's work groups engine. Each data type that derives from the livelong class has a field pointing to an associated parallel profile of the object. This parallel profile object is responsible for the data storage and actual tracking of physical location of the data pieces throughout the cluster's processes. The relationship between user type and parallel profile is regulated by means of providing a specialized constructor for the user's data type that sets the memory storage parameters according to the actual user's object parameters in a user defined way.

That is, one is free to design his own data types like parallel sparse matrix alongside with embedded parallel dense matrix type. The simpliest example is dense matrix - inside the specialized constructor we simply set the dim.x and dim.y of parallel profile to be matrix cols and matrix rows numbers respectively. This setting immidiately sets work group grid of the object considering also provided size of the object's work group. Alongside with the dimension other parameters are being set - such as data element size and work block/work item dimensions (the latter can be changed individually for the object during the runtime).

>>>>

In order to turn Ambient into a general purpose framework, we have decoupled the Ambient data representation from the user datatypes. There are two aspects to the parallel datatypes:

First is the object's lifetime. As we employ the technique of accumulated execution, we have to ensure that the objects that were created on the stack are still alive at the time of the actual execution. For this purpose we introduced \texttt { livelong } class that forces derived data types to duplicate themselves in heap memory upon construction if they were allocated on stack. Using this duplicate we can ensure that the object will stay alive until the moment of execution of the actual kernels using this data.

The second aspect is the actual connection to Ambient's work groups engine. Each data type derived from the livelong class has a field pointing to an associated parallel profile of the object. This parallel profile object is responsible for the data storage and actual tracking of the physical location of the data pieces throughout the cluster's processes. The relationship between user type and parallel profile is regulated by means of providing a specialized constructor for the user's data type that sets the memory storage parameters according to the actual user's object parameters in a user defined way.

That is, one is free to design his own data types like e.g. a parallel sparse matrix type, either alongside the embedded parallel dense matrix type, or to replace it if the user chooses to. The simpliest example would be for a dense matrix. Inside the specialized constructor we simply set the \texttt{dim.x} and \texttt{dim.y} of the parallel profile to be the numbers of columns and rows respectively. Assuming the size of the work group is provided, this immidiately sets the work group grid of the object. Alongside with the dimension other parameters are being set - such as the data element size and the work block/work item dimensions (the latter can be changed individually for the object during runtime).
